Chemical composition analysis for X-ray transport container scans. \\ Анализ химического состава при сканировании транспортных контейнеров гамма-излучением

### 2018-10-18 EMIN-2018/article/emin2108.tex
save

\label{sec:intro}
\section*{Introduction}
\newpage
\section*{Dual energy method}

@@

            $$
T(E_0, t, Z) = \frac{\int \limits_0^{E_0} S(E_0, E) \exp(-\mu(E,Z)\times t)~dE)}{\int \limits_0^{E_0} S(E_0, E)~dE}
$$
            $T$ -  transmittance, $S(E_0, E)$ - response function\\
$\mu(E,Z)$ - attenuation\\
$t$ -  optical thickness\\
$E_0$ -  up-limit energy of bremsstrahlung\\
$E$ - energy of gamma ray\\
$Z$ - charge of nuclei 
        \begin{itemize}
    \item We can't defined Z if optical thickness $t$ is unknown.
    \item But we can use two electron beams with different energy which give gamma rays with up-limit energy $E^{(1)}_0$ and $E^{(2)}_0$.
    \item Then we can get Z as a result of minimizing this function:
    $$
    F(z) = \frac{|t(E^{(1)}_0,z) - t(E^{(2)}_0,z)|}{t(E^{(1)}_0,z)} \to min
    $$
    \item This technique allows to determine scan object as a one from four possible groups: $Z_{eff} \sim 5$, $Z_{eff} \sim 13$, $Z_{eff} \sim 26$, $Z_{eff} \sim 82$.
\end{itemize}
Disadvantages of dual energy method
    \begin{itemize}
        \item It is too difficult to irradiate the target with beams with different energy.
        \item Low efficiency for target which contains elements with strongly different charges.
    \end{itemize}
Our method
    \begin{itemize}
        \item Use only one electron beam with energy $E = 10 MeV$.
        \item Measure not only the space distribution, but also the energy of gamma rays.
    \end{itemize}

@@

\section*{Thickness reconstruction}

\begin{figure}[t]
    \begin{center}
        % Сделать подписи к отдельным картинкам
        \includegraphics[width=120mm]{figures/yed_schema_2.pdf}  
        \vspace{-3mm}
        \caption{a) }
    \end{center}
    \labelf{fig01}
    \vspace{-5mm}
\end{figure}
                Attenuation of gamma ray flux is defined as:
$$
\frac{N(E)}{N_0(E)} = \exp(-\sum_i \Sigma^{mean}_i(E)x_i)
$$
where $x_i$ --- thickness of the $i$-layer, $\Sigma^{mean}_i$ --- mean cross-section for group of materials, $N,~N_0$ --- the number of gamma.
\begin{itemize}
    \item Disregard secondary scattering.
    \item Disregard the annihilation line.
\end{itemize}

                    \begin{itemize}
    \item Full thickness is known.
    \item Find thickness using least squares:

\end{itemize}
$$
\sum_E(\ln \frac{N(E)}{N_0(E)} + \sum_i \Sigma^{mean}_i(E)x_i))^2 \to min
$$

\begin{figure}[t]
    \begin{center}
        % Сделать подписи к отдельным картинкам
        \includegraphics[width=60mm]{figures/reconstruction.pdf} 
        \includegraphics[width=60mm]{figures/relError.pdf}  
        \vspace{-3mm}
        \caption{a) b)Distribution of reconstruction errors for various numerical experiments}
    \end{center}
    \labelf{fig01}
    \vspace{-5mm}
\end{figure}

\section*{Experiment}
The experiment was conducted by Dr. Guber and Dr. Ivashkin.
The sum of signals from 2 photodiode\\
Noise threshold: 100 KeV\\
\begin{center}
    \begin{tabular}[c]{|c|c|c|}
    \hline 
    Source & Energy, MeV & Sigma/Mean \\
    \hline 
    $^{22}Na$&0.511 & 14.7\%  \\ 
    \hline 
    $^{137}Cs$&0.662 & 19\%\\ 
    \hline 
    $^{22}Na$& 1.275 & 13\% \\
    \hline 
\end{tabular} 
\end{center}

\begin{figure}[t]
    \begin{center}
        % Сделать подписи к отдельным картинкам
        \includegraphics[width=80mm]{figures/yed.pdf}  
        \vspace{-3mm}
        \caption{a) }
    \end{center}
    \labelf{fig01}
    \vspace{-5mm}
\end{figure}
\section*{Conclusion}
% Выводы сделать более твердыми и акцентирванными
Results:   
    \begin{enumerate}
        \item The measurement of the gamma ray spectrum allows to identify cargo belonging to the group of materials with certain $Z_{eff}$.
        \item Also it allows to define the thickness of layers from different elements with the accuracy about 25\%.
        \item The energy resolution of the detector based on a BGO scintillator was studied.  For the photodetector with full array of pixels the energy resolution is expected about 10\%.
    \end{enumerate}
Plans and perspectives
    With financial support can be developed:
    \begin{enumerate}
        \item The program which checks cargo of a transport container  for compliance cargo manifest
        \item The algorithm for the 3D gamma-tomography.
    \end{enumerate}

@@

%\nocite{*}
\bibliographystyle{pepan}
\bibliography{pepan_biblio}

>>>>

\newpage
\label{sec:intro}
\section*{Introduction}
Geant4~\cite{ALLISON2016186, spirin}
\section*{Dual energy method}

@@

            $$
T(E_0, t, Z) = \frac{\int \limits_0^{E_0} S(E_0, E) \exp(-\mu(E,Z)\times t)~dE)}{\int \limits_0^{E_0} S(E_0, E)~dE},
$$
where
$T$ -  transmittance,
$S(E_0, E)$ - response function,
$\mu(E,Z)$ - attenuation,
$t$ -  optical thickness,
$E_0$ -  up-limit energy of bremsstrahlung,
$E$ - energy of gamma ray,
$Z$ - charge of nuclei.
We can't defined Z if optical thickness $t$ is unknown.
But we can use two electron beams with different energy which give gamma rays with up-limit energy $E^{(1)}_0$ and $E^{(2)}_0$.
 Then we can get Z as a result of minimizing this function:
    $$
    F(z) = \frac{|t(E^{(1)}_0,z) - t(E^{(2)}_0,z)|}{t(E^{(1)}_0,z)} \to min
    $$
This technique allows to determine scan object as a one from four possible groups: $Z_{eff} \sim 5$, $Z_{eff} \sim 13$, $Z_{eff} \sim 26$, $Z_{eff} \sim 82$.\\
Disadvantages of dual energy method
    \begin{itemize}
        \item It is too difficult to irradiate the target with beams with different energy.
        \item Low efficiency for target which contains elements with strongly different charges.
    \end{itemize}
Our method
    \begin{itemize}
        \item Use only one electron beam with energy $E = 10 MeV$.
        \item Measure not only the space distribution, but also the energy of gamma rays.
    \end{itemize}

@@

\begin{figure}[t]
    \begin{center}
        % Сделать подписи к отдельным картинкам
        % Переделать картинку от Туринге
        \includegraphics[width=60mm]{figures/UranCube1.pdf} 
        \includegraphics[width=60mm]{figures/UranCube2.pdf}  
        \vspace{-3mm}
        \caption{a) b)}
    \end{center}
    \labelf{fig01}
    \vspace{-5mm}
\end{figure}

\section*{Thickness reconstruction}
\begin{wrapfigure}[8]{r}{0.6\linewidth} 
    % Сделать подписи к отдельным картинкам
    \includegraphics[width=\linewidth]{figures/yed_schema_2.pdf}
    \vspace{-3mm}
    \caption{ Thickness reconstruction}
    \labelf{schema2}
    \vspace{-5mm}
\end{wrapfigure}
Мы будем рассматривать одномерный случай, когда гамма-кванты проходят скозь стопку из нескольких материалов с фиксированной полной толщиной и нам нужно восстановить толщины отдельных материалов (см. схему ~\ref{scema2}). Мы будем использовать простую физическую модель в которой ослабление задается  следующей формулой: 
\begin{equation}
\label{eq:gamma}
\frac{N(E)}{N_0(E)} = \exp(-\sum_i \Sigma^{mean}_i(E)x_i),
\end{equation}
where $x_i$ --- thickness of the $i$-layer, $\Sigma^{mean}_i$ --- mean cross-section for group of materials with близкими зарядами ядра, $N,~N_0$ --- the number of gamma.При этом мы не будем учитывать множественное рассеяние и наличие аннигиляционной линии. Мы считаем что полная толщина известна и для восстановления толщин отдельных слоев мы будем использовать метод наименьших квадратов, то есть минимизировать такую сумму:
\begin{equation}
\sum_E(\ln \frac{N(E)}{N_0(E)} + \sum_i \Sigma^{mean}_i(E)x_i))^2 \to min
\end{equation}
Далее приведен пример работы алгоритма. We consider the object which has 3 layers of aluminium, iron and lead.The figure ~\ref{rec:ex} shows a contribution of every reconstructed material in summary attenuation. The table ~\ref{tab:rec} contains the results of reconstruction.
\begin{figure}[t]
    \begin{center}
        % Сделать подписи к отдельным картинкам
        \includegraphics[width=60mm]{figures/reconstruction.pdf} 
        \includegraphics[width=60mm]{figures/relError.pdf}  
        \vspace{-3mm}
        \caption{a) b)Distribution of reconstruction errors for various numerical experiments}
    \end{center}
    \labelf{rec:ex}
    \vspace{-5mm}
\end{figure}

\begin{table}
\begin{center}
        \begin{tabular}[c]{|c|c|c|}
        \hline 
        Material & Real, cm & Reconstructed, cm \\ 
        \hline 
        Al & 20 & 19.6 \\ 
        \hline 
        Fe & 40 & 41.6 \\ 
        \hline 
        Pb & 30 & 28.7 \\ 
        \hline 
    \end{tabular}
\end{center}
\caption{Example of thickness reconstruction}
    \label{tab:rec}
\end{table}

\section*{Experiment}
\begin{wrapfigure}[20]{r}{0.5\linewidth} 
    % Сделать подписи к отдельным картинкам
    \includegraphics[width=\linewidth]{figures/yed.pdf}  
    \vspace{-3mm}
    \caption{Schema of experiment}
    \labelf{pic:experiment}
    \vspace{-5mm}
\end{wrapfigure}
Кроме моделирования, доктором Губером и доктором Ивашкины, проведены измерения энергического разрешения детектора  для регистрации гамма излучения. Схема эксперимента приведена на рисунке~\ref{pic:experiment}. В работе использовались сцинтиллятор BGO размером 10х30х100мм и фотодетектор ArrayC-60035-4P. Проводились измерения спектров от источников $\beta$– излучения (Na-22 и Cs-137). Энергии излучения 0.511 МэВ и 1.275 МэВ для Na и 0.662 МэВ для Cs. В качестве фотодетектора использовалась матрица из четырех фотодиодов ArrayC-60035-4P с размером фотодиода 6х6мм, снабженная индивидуальным предусилителем с коэффициентом усиления 150. В процессе работы измерялся суммарный сигнал с двух фотодиодов матрицы. Сигнал c матрицы подавался на усилитель (ORTEC 579) и затем поступал на входной канал АЦП (CAEN DT5742) и на дискриминатор (CAEN mod.224), логический сигнал которого служил триггером в системе. В отсутствие источника шкала АЦП была прокалибрована в абсолютных единицах – числе фотоэлектронов. Результат измерения энергитического разрешения приведен в таблице ~\ref{tab:ex}

The experiment was conducted by Dr. Guber and Dr. Ivashkin.
The sum of signals from 2 photodiode\\
Noise threshold: 100 KeV\\
\begin{table}
    \begin{center} 
        \begin{tabular}[c]{|c|c|c|}

            \hline 
            Source & Energy, MeV & Sigma/Mean \\
            \hline 
            $^{22}Na$&0.511 & 14.7\%  \\ 
            \hline 
            $^{137}Cs$&0.662 & 19\%\\ 
            \hline 
            $^{22}Na$& 1.275 & 13\% \\
            \hline 
        \end{tabular} 
    \end{center}
\caption{Energy resolution of detector}
\label{tab:ex}
\end{table}

\section*{Conclusion}
% Выводы сделать более твердыми и акцентирванными
Results:   
    \begin{enumerate}
        \item The measurement of the gamma ray spectrum allows to identify cargo belonging to the group of materials with certain $Z_{eff}$.
        \item Also it allows to define the thickness of layers from different elements with the accuracy about 25\%.
        \item The energy resolution of the detector based on a BGO scintillator was studied.  For the photodetector with full array of pixels the energy resolution is expected about 10\%.
    \end{enumerate}
Plans and perspectives
    With financial support can be developed:
    \begin{enumerate}
        \item The program which checks cargo of a transport container  for compliance cargo manifest
        \item The algorithm for the 3D gamma-tomography.
    \end{enumerate}

@@

%\nocite{*}
\bibliographystyle{pepan}
\bibliography{references.bib}
